---
title: "Replication project - 3 - Univariate analysis - T2DM - SGSS"
output:
  html_document: default
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "output") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(furniture)
library(knitr)
library(here)

# Recommended best practice - clears all values from the environment
rm(list = ls())
```

```{r load data, echo=FALSE}
t2dm_df <- readRDS(file=here("data", "processed-t2d_df_sgss_eth.rds"))

```

## Logistic regression models
### Factors associated with 28 day admission in diabetic patients
Here we use logistic regression to investigate possible factors associated with 28day hospital admission following COVID diagnosis in T2D patients Note, this does not include the controls (only those with diabetes).
Univariate logistic regression was performed including each candidate predictor in the model as the sole predictor.
Below is a table of the results presented as odds ratios and 95% confidence interval for each of the univariate models.

***NB this is using the SGSS data to determine COVID tests***

```{r , echo=FALSE, warning=FALSE, message = FALSE}

t2dm_df1 <- t2dm_df %>% dplyr::filter(diab==2)

fit <- glm(admission ~ Sex, data = t2dm_df1, family = "binomial")
diab2_coefs_int <- cbind(OR = exp(coef(fit)), exp(confint(fit)))[1,]
diab2_coefs_var <- cbind(OR = exp(coef(fit)), exp(confint(fit)))[2,]

#TODO don't hard code these
a <- names(t2dm_df1)[c(8,17:21,24,29:36,44)]

for(i in seq_along(a)){
  print(a[i])

  fit <- glm(admission ~ get(a[i]), data = t2dm_df1, family = "binomial")
  diab2_coefs_int <- rbind(diab2_coefs_int, cbind(OR = exp(coef(fit)), exp(confint(fit)))[1,] )
  diab2_coefs_var <- rbind(diab2_coefs_var, cbind(OR = exp(coef(fit)), exp(confint(fit)))[2,] )
}
b <- names(t2dm_df1[c(9,28,46)])

# townsend
fit <- glm(admission ~ as.factor(TownsendQuintileHigherIsMoreDeprived), data = t2dm_df1, family = "binomial")
diab2_coefs_int <- rbind(diab2_coefs_int, cbind(OR = exp(coef(fit)), exp(confint(fit)))[1,] )
diab2_coefs_var <- rbind(diab2_coefs_var, cbind(OR = exp(coef(fit)), exp(confint(fit)))[-1,] )

# smoking
fit <- glm(admission ~ as.factor(CurrentSmokingStatus), data = t2dm_df1, family = "binomial")
diab2_coefs_int <- rbind(diab2_coefs_int, cbind(OR = exp(coef(fit)), exp(confint(fit)))[1,] )
diab2_coefs_var <- rbind(diab2_coefs_var, cbind(OR = exp(coef(fit)), exp(confint(fit)))[-1,] )

# ethnicity
fit <- glm(admission ~ ethnicity, data = t2dm_df1, family = "binomial")
diab2_coefs_int <- rbind(diab2_coefs_int, cbind(OR = exp(coef(fit)), exp(confint(fit)))[1,] )
diab2_coefs_var <- rbind(diab2_coefs_var, cbind(OR = exp(coef(fit)), exp(confint(fit)))[-1,] )

b1 <- c(levels(as.factor(t2dm_df1$TownsendQuintileHigherIsMoreDeprived))[-1], levels(as.factor(t2dm_df1$CurrentSmokingStatus))[-1], levels(t2dm_df1$ethnicity)[-1] )

diab2_coefs_int <- data.frame(var= c("Sex", a, b), diab2_coefs_int)
diab2_coefs_var <- data.frame(var= c("Sex", a, b1), diab2_coefs_var)


```


```{r , echo=FALSE, warning=FALSE, message = FALSE}

kable(diab2_coefs_var, digits = 3, row.names = FALSE, align = "lccc", caption = "Odds ratios from logistic regression for 28day hospital admission following covid diagnosis in T2D patients including each variable in a univariable analysis" )

```

