---
title: "Replication project - 4 - Logistic Regression - T1DM vs controls - SGSS"
output:
  html_document: default
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "output") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(furniture)
library(knitr)
library(here)

# Recommended best practice - clears all values from the environment
rm(list = ls())
```

```{r load data, echo=FALSE}
t1dm_df <- readRDS(file=here("data", "processed-t1d_df_sgss_eth.rds"))
t1dm_df1 <- t1dm_df %>% dplyr::filter(diab==1)
```

## Logistic regression models
### Factors associated with 28 day admission in T1DM patients compared to non-diabetes patients
Here we use logistic regression to investigate possible factors associated with 28day hospital admission following COVID diagnosis in diabetes patients compared to non-diabetes patients. Note, this includes the controls.
Initially a logistic regression was performed in the T1DM cohort (including controls). The model included age, sex and diabetes. This provided an odds ratio for diabetes vs non-diabetes patients and was then used as the base model to compare other models.

Following this, we added BMI into the model and compared the odds ratio for diabetes to the base model. This was repeated for all of the variables of interest.

***NB this is using the SGSS data to determine COVID tests***

```{r , echo=FALSE, warning=FALSE, message = FALSE}

# odds ratio for T1D vs controls 
fit <- glm(admission ~ age + Sex + as.factor(diab), data = t1dm_df, family = "binomial")
test <- as.data.frame(cbind(OR = exp(coef(fit)), exp(confint(fit))))

```

```{r , echo=FALSE, warning=FALSE, message = FALSE}
# odds ratio for T1D vs controls 
fit <- glm(admission ~ age + Sex + as.factor(diab), data = t1dm_df, family = "binomial")

diab1_coef_diab <- cbind(OR = exp(coef(fit))[4], exp(confint(fit)))[4,]

a <- names(t1dm_df1)[c(8,17:21,24,29:36)]
a <- a[-9]
b <- names(t1dm_df1[c(9,28,46)])
for(i in seq_along(a)){

fit <- glm(admission ~ age + Sex + as.factor(diab) + get(a[i]), data = t1dm_df, family = "binomial")
diab1_coef_diab <- rbind(diab1_coef_diab, cbind(OR = exp(coef(fit))[4], exp(confint(fit)))[4,] )

}
fit <- glm(admission ~ age + Sex + as.factor(diab) + as.factor(TownsendQuintileHigherIsMoreDeprived), data = t1dm_df, family = "binomial")
diab1_coef_diab <- rbind(diab1_coef_diab, cbind(OR = exp(coef(fit))[4], exp(confint(fit)))[4,] )

fit <- glm(admission ~ age + Sex + as.factor(diab) + as.factor(CurrentSmokingStatus), data = t1dm_df, family = "binomial")
diab1_coef_diab <- rbind(diab1_coef_diab, cbind(OR = exp(coef(fit))[4], exp(confint(fit)))[4,] )

fit <- glm(admission ~ age + Sex + as.factor(diab) + as.factor(ethnicity), data = t1dm_df, family = "binomial")
diab1_coef_diab <- rbind(diab1_coef_diab, cbind(OR = exp(coef(fit))[4], exp(confint(fit)))[4,] )

diab1_coef_diab <- data.frame(var = c("base", a, b), diab1_coef_diab)

```

```{r , echo=FALSE, warning=FALSE, message = FALSE}
kable(diab1_coef_diab, digits = 3, row.names = FALSE, align = "lccc", caption = "Diabetes coefficient from each logistic model in T1Ds matched cohort - with SGSS data for COVID tests" )

```

```{r , echo=FALSE, warning=FALSE, message = FALSE}
library(ggplot2)

diab1_coef_diab$index <- 1:nrow(diab1_coef_diab)

g <- ggplot(data=diab1_coef_diab, aes(y=index, x=OR, xmin = X2.5.. , xmax = X97.5..)) +
  geom_point() +
  geom_errorbarh(height=.1) +
  scale_y_continuous(breaks=1:nrow(diab1_coef_diab), labels = diab1_coef_diab$var) +
  coord_cartesian(xlim= c(0.5,4)) +
  geom_vline(xintercept = 1, color = 'black', linetype='dashed', alpha=.5) +
  labs(title = "T1D coefficient from each model - using SGSS data for COVID tests", x='Odds ratio', y = 'Variable') +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5, size=11))

print(g)

```

