---
title: "Replication project - 4 - Logistic Regression - T2DM vs controls - SGSS"
output:
  html_document: default
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "output") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(furniture)
library(knitr)
library(here)

# Recommended best practice - clears all values from the environment
rm(list = ls())
```

```{r load data, echo=FALSE}
t2dm_df <- readRDS(file=here("data", "processed-t2d_df_sgss_eth.rds"))
t2dm_df2 <- t2dm_df %>% dplyr::filter(diab==2)
```

## Logistic regression models
### Factors associated with 28 day admission in T2DM patients compared to non-diabetes patients
Here we use logistic regression to investigate possible factors associated with 28day hospital admission following COVID diagnosis in diabetes patients compared to non-diabetes patients. Note, this includes the controls.
Initially a logistic regression was performed in the T2DM cohort (including controls). The model included age, sex and diabetes. This provided an odds ratio for diabetes vs non-diabetes patients and was then used as the base model to compare other models.

Following this, we added BMI into the model and compared the odds ratio for diabetes to the base model. This was repeated for all of the variables of interest.

***NB this is using the SGSS data to determine COVID tests***

```{r , echo=FALSE, warning=FALSE, message = FALSE}
# odds ratio for T2D vs controls 
fit <- glm(admission ~ age + Sex + as.factor(diab), data = t2dm_df, family = "binomial")

diab2_coef_diab <- cbind(OR = exp(coef(fit))[4], exp(confint(fit)))[4,]

a <- names(t2dm_df2)[c(8,17:21,24,29:36)]
b <- names(t2dm_df2[c(9,28,46)])
for(i in seq_along(a)){

fit <- glm(admission ~ age + Sex + as.factor(diab) + get(a[i]), data = t2dm_df, family = "binomial")
diab2_coef_diab <- rbind(diab2_coef_diab, cbind(OR = exp(coef(fit))[4], exp(confint(fit)))[4,] )

}
fit <- glm(admission ~ age + Sex + as.factor(diab) + as.factor(TownsendQuintileHigherIsMoreDeprived), data = t2dm_df, family = "binomial")
diab2_coef_diab <- rbind(diab2_coef_diab, cbind(OR = exp(coef(fit))[4], exp(confint(fit)))[4,] )

fit <- glm(admission ~ age + Sex + as.factor(diab) + as.factor(CurrentSmokingStatus), data = t2dm_df, family = "binomial")
diab2_coef_diab <- rbind(diab2_coef_diab, cbind(OR = exp(coef(fit))[4], exp(confint(fit)))[4,] )

fit <- glm(admission ~ age + Sex + as.factor(diab) + as.factor(ethnicity), data = t2dm_df, family = "binomial")
diab2_coef_diab <- rbind(diab2_coef_diab, cbind(OR = exp(coef(fit))[4], exp(confint(fit)))[4,] )

diab2_coef_diab <- data.frame(var = c("base", a, b), diab2_coef_diab)

```

```{r , echo=FALSE, warning=FALSE, message = FALSE}
kable(diab2_coef_diab, digits = 3, row.names = FALSE, align = "lccc", caption = "Diabetes coefficient from each logistic model in T2Ds matched cohort - with SGSS data for COVID tests" )

```

```{r , echo=FALSE, warning=FALSE, message = FALSE}
library(ggplot2)

diab2_coef_diab$index <- 1:nrow(diab2_coef_diab)

g <- ggplot(data=diab2_coef_diab, aes(y=index, x=OR, xmin = X2.5.. , xmax = X97.5..)) +
  geom_point() +
  geom_errorbarh(height=.1) +
  scale_y_continuous(breaks=1:nrow(diab2_coef_diab), labels = diab2_coef_diab$var) +
  coord_cartesian(xlim= c(0.6,2.5)) +
  geom_vline(xintercept = 1, color = 'black', linetype='dashed', alpha=.5) +
  labs(title = "T2D coefficient from each model - using SGSS data for COVID tests", x='Odds ratio', y = 'Variable') +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5, size=11))

print(g)
```
